Title

### 2024-06-09 paper/main.tex
Updates

%File: anonymous-submission-latex-2024.tex
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage[submission]{aaai24}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet}  % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in} % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in} % DO NOT CHANGE THIS
%
% These are recommended to typeset algorithms but not required. See the subsubsection on algorithms. Remove them if you don't have algorithms in your paper.
\usepackage{algorithm}
\usepackage{algorithmic}

%
% These are are recommended to typeset listings but not required. See the subsubsection on listing. Remove this block if you don't have listings in your paper.
\usepackage{newfloat}
\usepackage{listings}
\DeclareCaptionStyle{ruled}{labelfont=normalfont,labelsep=colon,strut=off} % DO NOT CHANGE THIS
\lstset{%
	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
	aboveskip=0pt,belowskip=0pt,%
	showstringspaces=false,tabsize=2,breaklines=true}
\floatstyle{ruled}
\newfloat{listing}{tb}{lst}{}
\floatname{listing}{Listing}
%
% Keep the \pdfinfo as shown here. There's no need
% for you to add the /Title and /Author tags.
\pdfinfo{
/TemplateVersion (2024.1)
}

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{todonotes}
\usepackage{bm}
\usepackage{subcaption}
% \usepackage[ruled]{algorithm2e}

@@

% The file aaai24.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%

% Title

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Title}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

    % Sunil Issar\textsuperscript{\rm 2},
    % J. Scott Penberthy\textsuperscript{\rm 3},
    % George Ferguson\textsuperscript{\rm 4},
    % Hans Guesgen\textsuperscript{\rm 5}
    % Note that the comma should be placed after the superscript

    1900 Embarcadero Road, Suite 101\\
    Palo Alto, California 94303-3310 USA\\
    % email address must be in roman text type, not monospace or sans serif
    proceedings-questions@aaai.org
%
% See more examples next
}

%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
\iffalse
\title{My Publication Title --- Single Author}
\author {
    Author Name
}
\affiliations{
    Affiliation\\
    Affiliation Line 2\\
    name@example.com
}
\fi

\iffalse
%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
\title{My Publication Title --- Multiple Authors}
\author {
    % Authors
    First Author Name\textsuperscript{\rm 1},
    Second Author Name\textsuperscript{\rm 2},
    Third Author Name\textsuperscript{\rm 1}
}
\affiliations {
    % Affiliations
    \textsuperscript{\rm 1}Affiliation 1\\
    \textsuperscript{\rm 2}Affiliation 2\\
    firstAuthor@affiliation1.com, secondAuthor@affilation2.com, thirdAuthor@affiliation1.com
}
\fi

@@

\begin{abstract}
	The initial step in a causal analysis involves determining the causal
	structure between the variables from data typically in the form of a
	Directed Acyclic Graph (DAG) or an Structural Equation Model (SEM).
	This procedure, known as \emph{causal discovery}, has been extensively
	studied for both DAGs and SEMs. While numerous automated algorithms
	exist for estimating DAGs from datasets, their application in applied
	fields remain limited, possibly due to a lack of confidence in these
	algorithms stemming from them making obvious mistakes, and the
	difficulties in choosing the right algorithm for a given dataset.
	Consequently, in applied disciplines, these DAGs are predominantly
	constructed manually based on domain expertise. Due to being
	constructed by hand, it is vital that the fit of these models are
	tested against data and the models are potentially modified to acheive
	a better fit. On the other hand, in the field of SEMs models are
	typically constructed manually through an iterative process of
	assessing fit and modifying the model according to that. This process
	is commonly known as \emph{Specification Search} where methods like
	\emph{modification indices} combined with expert knowledge can be used
	to guide this modification process. In the case of DAGs, although there
	are tests to assess the global fit of a model by combining the tests
	using implied conditional independencies of the model, but there are no
	methods that can guide our modification process. In this paper, we
	present a simple modification process aimed at helping researchers in
	manually constructing or modifying DAGs outputted by an algorithm. This
	modification process is based on using a (conditional) measure of
	association between variables in the model to assess potential edges
	that best explains unexplained correlation between variables which the
	researcher specifies the direction of this edge. In the case of
	continuous or discrete variables, the effect size measures of local
	tests can be used as this measure of association, however for mixed
	data there is no such measure. We present a measure of association
	based on canonical correlations for mixed data. We also present a
	graphical web-tool that can assist researchers in iteratively modifying
	and constructing their DAGs. We theoretically show that in the presence
	of an oracle, this iterative modification process is consistent in
	recovering the true DAG. Emprically, we show that using the mixed data
	measure of association, the modification process performs comparably to
	PC and Hill-Climb Search algorithms if the user is able to correctly
	identify the direction of the edge in one out of three cases.
\end{abstract}

@@

\emph{Causal discovery} is the process of identifying and understanding causal
relationships among variables in a dataset.

To perform any causal effect identification or estimation using Directed
Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
first construct a causal structure. This process of constructing the casual
structure from data is known as \emph{causal discovery}. In the field of DAGs,
a lot of effort has been put on developing automated algorithms that can
construct these DAGs from data. These algorithm takes different approaches to
causal discovery such as Constraint-based algorithms such as PC and Fast Causal
Infenrece (FCI) where the algorithm tries to match the conditional
independences in the data to the ones impled by the model, score-based methods
such as Hill-Climb Search, Greedy Equivalence Search (GES)  where the algorithm
tries to find a model that has the best score given some scoring metric, and
continuous optimization where the problem is formulated as a constrained
optimization problem such as No Tears.

@@

\begin{enumerate}
	\item Biased towards sparse graphs.
	\item Algorithms making obvious mistakes: These algorithms often make
		very obvious mistakes, making it harder for the user to trust
		them.
	\item Difficulties in choosing an algorithm: For any given dataset, the
		user need to make a choice of the algorithm. Given that these
		algorithms work very differently, it can be difficult to make a
		choice. Additionally, each of these algorithms have
		hyperparameters and choices to make for them, making it
		difficult. As there is no way to evaluate the performance of
		these algorithms on our dataset.
\end{enumerate}

@@

We present a similar procedure for modifying DAGs based on expert knowledge and
measure of association. We first use the measure of association to test whether
the association observed in the data is explained by the model structure. Based 
on this testing, we suggest adding edges by ranking them by the most unexplained
association between variables. The researcher can then use expert knowledge to 
decide the direction of the edge between the variables.

@@

\begin{figure}
		\centering
		\begin{tikzpicture}
			\begin{scope}[xshift=-2.5cm, yshift=0.7cm, scale=0.7]
			\tikzstyle{every node}=[align=center, inner sep=1pt]
				\node (sex) at (-0.7, -0.8) {Sex};
				\node (age) at (-0.3, 0) {Age};
				\node (ed) at (-1.8, -1.6) {Education};
				\node (occ) at (1.8, 0) {Occupation};
				\node (hrpw) at (1.3, -1.6) {Hours Per \\ Week};
				\node (income) at (-0.7, -2.8) {Income};

				\draw[->]  (age) to[bend right=20] (ed);
				\draw[->]  (sex) to (ed);
				\draw[->]  (age) to (hrpw);
				\draw[->]  (ed) to (hrpw);
				\draw[->]  (sex) to (hrpw);
				\draw[->]  (ed) to (income);
				\draw[->]  (hrpw) to (income);
				\draw[->]  (occ) to[out=300, in=0, looseness=1.4] (income.east);
				\draw[->]  (sex) to (income);
				\draw[->]  (ed) to[out=90, in=150, looseness=1.3] (occ);
				\draw[->]  (sex) to (occ);	
			\end{scope}
			\begin{scope}[xshift=-3cm]
				\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
				\node[rectangle, align=center, inner sep=1pt] at (5, 0) {
					\begin{minipage}{.2\textwidth}
						\begin{equation*}
							\begin{split}
								\textnormal{Age} &\ci \textnormal{Incm} \rvert \textnormal{Edct, Hrpw, Sex} \\
								\textnormal{Age} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
								\textnormal{Age} &\ci \textnormal{Sex} \\
								\textnormal{Hrpw} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
							\end{split}
						\end{equation*}
					\end{minipage}
					};
			\end{scope}
		\end{tikzpicture}
		\caption{\todo[inline]{Show a full example of testing using this model}}
\end{figure}

@@

% \begin{algorithm}
% 	\caption{Expert Simulation}
% 	\While{True}{
% 	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $\;
% 	   $ e \gets e - \textnormal{blacklist} $\;
% 	   \If {$ e = 0 $}{
% 		$ G \gets G - e $ \;
% 		$ \textnormal{blacklist} \gets \textnormal{blacklist} \cup e $\;
% 	   } \;
% 	   \If{$ \textnormal{len}(e) = 0$}{
% 		   $ \textnormal{return}(G) $\;
% 	   }
% 	   $ X, Y \gets \max_{X, Y} e $\;
% 	   $ dir \gets Expert(X, Y, \alpha) $\;
% 	   $ G \gets G \cup dir $\;
% 	}
% 
% \end{algorithm}

@@

\subsection{Results}
\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics{../code/plots/shd_ribbon.pdf}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics{../code/plots/sid_ribbon.pdf}
	\end{subfigure}
\end{figure}

@@

\bibliography{references}
\end{document}

>>>>

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024

% ready for submission
\usepackage{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

% Packages added by me.
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{todonotes}
\usepackage{bm}
\usepackage{subcaption}
\usepackage[ruled]{algorithm2e}

@@

\title{Title}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

@@

\begin{abstract}
	In Pearl's framework, the initial step for performing a causal analysis
	involves determining the causal structure between the variables from
	data typically in the form of a Directed Acyclic Graph (DAG) or an
	Structural Equation Model (SEM). This procedure, known as \emph{causal
	discovery}, has been extensively studied for both DAGs and SEMs. While
	numerous automated algorithms exist for estimating DAGs from datasets,
	their application in applied fields remain limited, possibly due to a
	lack of confidence in these algorithms stemming from them making
	obvious mistakes, and the difficulties in choosing the right algorithm
	for a given dataset. Consequently, in applied disciplines, these DAGs
	are predominantly constructed manually based on domain expertise. Due
	to being constructed by hand, it is vital that the fit of these models
	are tested against data and the models are potentially modified to
	acheive a better fit. On the other hand, in the field of SEMs models
	are typically constructed manually through an iterative process of
	assessing fit and modifying the model according to that. This process
	is commonly known as \emph{Specification Search} where methods like
	\emph{modification indices} combined with expert knowledge can be used
	to guide this modification process. In the case of DAGs, although there
	are tests to assess the global fit of a model by combining the tests
	using implied conditional independencies of the model, but there are no
	methods that can guide our modification process. In this paper, we
	present a simple modification process aimed at helping researchers in
	manually constructing or modifying DAGs outputted by an algorithm. This
	modification process is based on using a (conditional) measure of
	association between variables in the model to assess potential edges
	that best explains unexplained correlation between variables which the
	researcher specifies the direction of this edge. In the case of
	continuous or discrete variables, the effect size measures of local
	tests can be used as this measure of association, however for mixed
	data there is no such measure. We present a measure of association
	based on canonical correlations for mixed data. We also present a
	graphical web-tool that can assist researchers in iteratively modifying
	and constructing their DAGs. We theoretically show that in the presence
	of an oracle, this iterative modification process is consistent in
	recovering the true DAG. Emprically, we show that using the mixed data
	measure of association, the modification process performs comparably to
	PC and Hill-Climb Search algorithms if the user is able to correctly
	identify the direction of the edge in one out of three cases.
\end{abstract}

@@

To perform any causal effect identification or estimation using Directed
Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
first construct a causal structure commonly known as DAG or path diagram. This
process of constructing the casual structure from data is known as \emph{Causal
Discovery}. In the field of DAGs, plenty of algorithms have been developed to
automatically construct DAGs from data. These algorithm takes different
approaches to causal discovery such as Constraint-based algorithms such as PC
and Fast Causal Infenrece (FCI) where the algorithm tries to match the
conditional independences in the data to the ones impled by the model,
score-based methods such as Hill-Climb Search, Greedy Equivalence Search (GES)
where the algorithm tries to find a model that has the best score given some
scoring metric, and continuous optimization where the problem is formulated as
a constrained optimization problem such as No Tears.

@@

\begin{enumerate}
	\item Can only give a Markov Equivalence class (MEC), and then experts
		need to rely on their domain expertise to orient these edges.
	\item Biased towards sparse graphs.
	\item Algorithms making obvious mistakes: These algorithms often make
		very obvious mistakes, making it harder for the user to trust
		them.
	\item Difficulties in choosing an algorithm: For any given dataset, the
		user need to make a choice of the algorithm. Given that these
		algorithms work very differently, it can be difficult to make a
		choice. Additionally, each of these algorithms have
		hyperparameters and choices to make for them, making it
		difficult. As there is no way to evaluate the performance of
		these algorithms on our dataset.
\end{enumerate}

@@

\todo[inline]{Other methods: Modification indices and Lagrange Multiplier tests to
selectively add edges And using Wald test or z statistics (also known as
critical ratios) for removing edges.}

One of the drawbacks of these global fit metrics is that they do not only test
for the structure but also the fitted parameterization.

We present a similar procedure for modifying DAGs based on expert knowledge and
measure of association. We first use the measure of association to test whether
the association observed in the data is explained by the model structure. 
We can not directly use the local CI tests to make modifications because that would give
way too many independencies and we have no way to rank them.
Based 
on this testing, we suggest adding edges by ranking them by the most unexplained
association between variables. The researcher can then use expert knowledge to 
decide the direction of the edge between the variables.

@@

\begin{figure}
		\centering
		\begin{tikzpicture}
			\begin{scope}[yshift=0.7cm, scale=1]
			\tikzstyle{every node}=[align=center, inner sep=1pt]
				\node (sex) at (-0.7, -0.8) {Sex};
				\node (age) at (-0.3, 0) {Age};
				\node (ed) at (-1.8, -1.6) {Education};
				\node (occ) at (1.8, 0) {Occupation};
				\node (hrpw) at (1.3, -1.6) {Hours Per \\ Week};
				\node (income) at (-0.7, -2.8) {Income};

				\draw[->]  (age) to[bend right=20] (ed);
				\draw[->]  (sex) to (ed);
				\draw[->]  (age) to (hrpw);
				\draw[->]  (ed) to (hrpw);
				\draw[->]  (sex) to (hrpw);
				\draw[->]  (ed) to (income);
				\draw[->]  (hrpw) to (income);
				\draw[->]  (occ) to[out=300, in=0, looseness=1.4] (income.east);
				\draw[->]  (sex) to (income);
				\draw[->]  (ed) to[out=90, in=150, looseness=1.3] (occ);
				\draw[->]  (sex) to (occ);	
			\end{scope}
			\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
			\node[rectangle, align=center, inner sep=1pt] at (6, 0) {
				\begin{minipage}{\textwidth}
					\begin{equation*}
						\begin{split}
							\textnormal{Age} &\ci \textnormal{Incm} \rvert \textnormal{Edct, Hrpw, Sex} \\
							\textnormal{Age} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
							\textnormal{Age} &\ci \textnormal{Sex} \\
							\textnormal{Hrpw} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
						\end{split}
					\end{equation*}
				\end{minipage}
				};
		\end{tikzpicture}
		\caption{\todo[inline]{Show a full example of testing using this model}}
\end{figure}

@@

\begin{algorithm}
	\caption{Expert Simulation}
	\While{True}{
	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $\;
	   $ e \gets e - \textnormal{blacklist} $\;
	   \If {$ e = 0 $}{
		$ G \gets G - e $ \;
		$ \textnormal{blacklist} \gets \textnormal{blacklist} \cup e $\;
	   } \;
	   \If{$ \textnormal{len}(e) = 0$}{
		   $ \textnormal{return}(G) $\;
	   }
	   $ X, Y \gets \max_{X, Y} e $\;
	   $ dir \gets Expert(X, Y, \alpha) $\;
	   $ G \gets G \cup dir $\;
	}

\end{algorithm}

@@

\subsection{Results}
\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics{../code/plots/shd_ribbon.pdf}
	\end{subfigure}%
	\begin{subfigure}{0.5\textwidth}
		\includegraphics{../code/plots/sid_ribbon.pdf}
	\end{subfigure}
\end{figure}

@@

\bibliographystyle{plain}
\bibliography{references.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Appendix / supplemental material}

Optionally include supplemental material (complete proofs, additional experiments and plots) in appendix.
All such materials \textbf{SHOULD be included in the main submission.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\include{checklist}

\end{document}
